A properly labelled proof system for intuitionistic modal logics

### 2019-09-16 caleidoscope/caleidoscope.tex
done section 5: completeness

>>>>

\begin{theorem}
	For any formula $A$. If $A$ is a theorem of $\IK$ then $A$ is provable in $\labIKp +\labrn{cut}$.
\end{theorem}

\begin{proof}
	In order to prove completeness, we present a syntactic completeness proof with respect the Hilbert system which means that we prove all Hilbert axioms using the rules from our system $\labIKp$: we give a proof for all the axioms of propositional intuionistic logic, for the five variants of $\mathsf{k}$ axiom from the intuitionistic syntax and finally, we simulate the necessitation rule and modus ponens.

	As we mentioned, in the course of the proof, we have to derive the five $\kax$ axioms. In the derivation of $\kax[1]$ is illustrated the need of having the rules corresponding to $\rn{refl}$ and $\rn{trans}$ in the system:

	\kone

	%%%
	We can also show that the rule $\rn{refl}$ is needed in the derivation of $\kax[2]$:

	\ktwo

	%%%%
	In order to continue our proof for the five variants of $\kax$, we display the proof for the axiom $\kax[3]$ which also illustrates the need of having the rule corresponding to $\rn{refl}$:

	\begin{center}

		$\vlderivation {
			\vlin{\rlabrn{\IMP}}
			{}
			{\Rightarrow \labels{x}{\DIA (A \OR B) \IMP (\DIA A \OR \DIA B)}}
			{\vlin {\llabrn{\DIA}}
				{}
				{x \le y, \labels{y}{\DIA (A \OR B)} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vliin{\llabrn{\OR}}{}{x \le y, yRz, \labels{z}{A \OR B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}{\vlhy{\circledast \hspace{40mm}}}{\vlhy{\star}}}}
		}$

	After applying the rule $\llabrn{\OR}$, our proof continues with two premises. On one side, the premise that follows from $\circledast$:

	\bigskip
	\begin{center}
		$\vlderivation{\vlin{\llabrn{\OR}}{}{\circledast}{\vlin {\rlabrn{\OR}}
				{}
				{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vlin {\rlabrn{\DIA}}
					{}
					{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
					{\vlin {\rn{refl}}
						{}
						{x \le y, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
						{\vlin {\rn{id}}
							{}
							{x \le y, z \le z, yRz, \labels{z}{A} \Rightarrow \labels{y}{\DIA A}, \labels{z}{A}, \labels{y}{\DIA B}}
							{\vlhy {}}}}}}}$
	\end{center}	
	And on the other hand, we have that the derivation continues from $\star$ like this:

	\begin{center}
		$\vlderivation{\vlin {\llabrn{\OR}}
			{}
			{\star}
			{\vlin {\rlabrn{\OR}}
				{}
				{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A \OR \DIA B}}
				{\vlin {\rlabrn{\DIA}}
					{}
					{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}}
					{\vlin {\rn{refl}}
						{}
						{x \le y, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
						{\vlin {\rn{id}}
							{}
							{x \le y, z \le z, yRz, \labels{z}{B} \Rightarrow \labels{y}{\DIA A}, \labels{y}{\DIA B}, \labels{z}{B}}
							{\vlhy {}}}}}}}$
	\end{center}
\end{proof}

The proof of $\kax[4]$ shows the need of having the rule corresponding to the frame conditions $\rn{F_1}$:

\kfour

To complete the proof of the five variants of $\kax$ we still have to see that we can prove $\kax[5]$:

\begin{center}

	$\vlderivation {
		\vlin{\rlabrn{\IMP}}
		{}
		{\Rightarrow \labels{x}{\DIA \BOT \IMP \BOT}}
		{\vlin {\llabrn{\DIA}}
			{}
			{x \le y, \labels{y}{\DIA \BOT} \Rightarrow \labels{y}{\BOT}}
			{\vlin {\llabrn{\BOT}}
				{}
				{x \le y, yRz, \labels{z}{\BOT} \Rightarrow \labels{y}{\BOT}}
				{\vlhy {}}}}
	}$

As we mentioned, in the course of the proof of the theorem, we have to give a proof for all the axioms of propositional intuitionistic logic. As an example, we display the proof of the axiom $\rn{AND}$-$\rn{2}$: $ \fm{A \AND B \IMP B}$:

\begin{center}
	$\vlderivation {
		\vlin{\rlabrn{\IMP}}
		{}
		{\Rightarrow \labels{x}{A \AND B \IMP B}}
		{\vlin {\llabrn{\AND}}
			{}
			{x \le y, \labels{y}{A \AND B} \Rightarrow \labels{y}{B}}
			{\vlin {\rn{refl}}
				{}
				{x \le y, \labels{y}{A}, \labels{y}{B} \Rightarrow \labels{y}{B}}
				{\vlin {\rn{id}}
					{}
					{x \le y, y \le y, \labels{y}{A}, \labels{y}{B} \Rightarrow \labels{y}{B}}
					{\vlhy {}}}}}
	}$
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
